Quantum Interval-Valued Probability: Contextuality and the Born Rule

### 2017-11-19 prop_letter.tex
prop_letter.tex Updating...

We also investigate the second key aspect of imprecise quantum
measurement, which is its statistical nature as reflected in the
application of the Born rule determining the probability of a given
measurement outcome~\citep{Born1983bibTeX,Mermin2007,Jaeger2007}.  Here,
Gleason's theorem establishes that there is no alternative to the Born
rule by demonstrating, using reasonable continuity arguments in
Hilbert spaces of dimension $d\ge3$, the existence of a unique state
$\rho$ consistent with the statistical predictions computed from the Born
rule~\citep{gleason1957,Redhead1987-REDINA,peres1995quantum}.  Our
question then concerns what happens when infinite precision cannot be
achieved.  We are able to show that, while a QIVPM incorporating the
effects of finite precision might not be consistent with Gleason's
unique state~$\rho$ on all projectors defined on a Hilbert space
$\Hilb$ of dimension $d\ge3$, there is a mathematically precise sense
in which one recovers the original Gleason theorem asymptotically.
Specifically, it is possible to construct a class of QIVPMs
representing bounded resources that is parameterized by the size of
the intervals. We then demonstrate that all
QIVPMs in this class are consistent with a non-empty set of quantum
states, and as the size of the intervals characterizing the
uncertainties goes to zero, this set of quantum states converges to
the unique state consistent with the Born rule. We furthermore
establish a mathematical correlation between the sizes of the
intervals and the volume of states that are all consistent with an
imprecise measurement.
\yutsung{``volume of states" are not defined in the Gleason's section.}

@@

\begin{definition}[$\delta$-Determinism]\label{def:delta-deterministic} A
  QIVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}$ is
  $\delta$-deterministic if, for every event $P\in\events$, we have
  that either 
  $\bar{\mu}\left(P\right)\subseteq\left[0,\delta\right]$ or
  $\bar{\mu}\left(P\right)\subseteq\left[1-\delta,1\right]$. 
\end{definition}
\yutsung{Clarify the domain of this.}

\begin{table}
  \noindent \centering{}\caption{\label{tab:classical-IVPMs}Example 
    QIVPMs~$\bar{\mu}_{A}$, $\bar{\mu}_{B}$ and $\bar{\mu}_{C}$ on
    commuting events $P$.}
\begin{tabular}{cccc}
\toprule 
\addlinespace
$P$ & $\bar{\mu}_{A}(P)$ & $\bar{\mu}_{B}(P)$ & $\bar{\mu}_{C}(P)$\tabularnewline\addlinespace
\midrule
\midrule 
\addlinespace
$\mathbb{0}$ & $\imposs$ & $\imposs$ & $\imposs$\tabularnewline\addlinespace
\midrule 
\addlinespace
$\proj{0}$ & $\necess$ & $\left[\frac{2}{3},1\right]$ & $\left[\frac{2}{3},\frac{2}{3}\right]$\tabularnewline\addlinespace
\midrule 
\addlinespace
$\proj{1}$ & $\imposs$ & $\left[0,\frac{1}{3}\right]$ & $\left[\frac{1}{3},\frac{1}{3}\right]$\tabularnewline\addlinespace
\midrule 
\addlinespace
$\mathbb{1}$ & $\necess$ & $\necess$ & $\necess$\tabularnewline\addlinespace
\bottomrule
\end{tabular}
\end{table}

@@

Our framework allows us to generalize the above theorem to state that
for small enough $\delta$, it is impossible to have
$\delta$-deterministic QIVPMs. We next prove a main result of this
paper which is a stronger statement of contextuality that includes the
effects of finite-precision.  Every QIVPM must map some events to
truly uncertain intervals, not just ``almost definite intervals.'' 
\yutsung{contextuality is never defined.}

@@

We have thus quantified one important aspect of uncertainty in quantum
mechanics---the effect of the imprecise nature of devices---which is a novel
addition to the theory of measurement. Indeed, as Heisenberg
emphasized in his famous microscope
example~\cite{Heisenberg1983apsrev4}, the conventional theory
of measurement states that it is impossible to precisely measure any
property of a system without disturbing it somewhat. Thus, there are
fundamental limits to what one can measure and these limits have traditionally 
been attributed to complementarity. Our imprecision represents an 
\emph{additional} source of indeterminacy beyond the inherent probabilistic nature
of quantum mechanics.
\yutsung{To Amr, the ``beyond'' is changed from ``to''. Andy wants to make sure
you are OK with it.}
% The significance of this observation is that the uncertainty in the
% intervals is, in contrast to the classical case, \emph{a fundamental
%   part of quantum theory}. There are in fact two sources of
% uncertainty: one due to the finite measurement resources and one due
% to complementarity, which requires every quantum framework to be truly
% probabilistic.

@@

Foundational concepts in quantum mechanics, such as the Kochen-Specker
and Gleason theorems, rely in subtle ways on the use of
unbounded resources. By assuming infinitely precise measurements,
these two insightful theorems form the foundations of two fundamental aspects of
quantum mechanics. On the one hand, the Kochen-Specker result reveals a
distinctive aspect of physical reality, the fact that it is
contextual, and, on the other hand, Gleason's theorem establishes a
relationship between quantum states and probabilities uniquely defined
by Born's rule.  Our goal in this paper has been to analyze the
physical consequences of a mathematical framework that allows for
finite precision measurements by introducing the concept of quantum
interval-valued probability. This framework incorporates uncertainty
in the measurement results by defining fuzzy probability measures, and
includes standard quantum measurement as the particular instance of
sharp, infinitely precise, intervals.  
\yutsung{``quantum measurement'' should be replaced by the }
In addition, we showed how
these two theorems emerge as limiting cases of this same framework,
thus connecting two seemingly unrelated aspects of quantum physics. We
noted that {\it arbitrary\/} finite precision measurement conditions
can nullify the main tenets of both theorems. However, by carefully
specifying experimental uncertainties, we were able to establish
rigorous bounds on the validity of these two theorems. Therefore, we
have established a context in which infinite precision quantum
mechanical theories can be reconciled with finite precision quantum
mechanical measurements, and have provided a possible resolution of
the Meyer-Mermin debate on the impact of finite precision on the
Kochen-Specker theorem~\cite{PhysRevLett.83.3751,Mermin1999}.

>>>>

We also investigate the second key aspect of imprecise quantum
measurement, which is its statistical nature as reflected in the
application of the Born rule determining the probability of a given
measurement outcome~\citep{Born1983bibTeX,Mermin2007,Jaeger2007}.  Here,
Gleason's theorem establishes that there is no alternative to the Born
rule by demonstrating, using reasonable continuity arguments in
Hilbert spaces of dimension $d\ge3$, the existence of a unique state
$\rho$ consistent with the statistical predictions computed from the Born
rule~\citep{gleason1957,Redhead1987-REDINA,peres1995quantum}.  Our
question then concerns what happens when infinite precision cannot be
achieved.  We are able to show that, while a QIVPM incorporating the
effects of finite precision might not be consistent with Gleason's
unique state~$\rho$ on all projectors defined on a Hilbert space
$\Hilb$ of dimension $d\ge3$, there is a mathematically precise sense
in which one recovers the original Gleason theorem asymptotically.
Specifically, it is possible to construct a class of QIVPMs
representing bounded resources that is parameterized by the size of
the intervals. We then demonstrate that all
QIVPMs in this class are consistent with a non-empty set of quantum
states, and as the size of the intervals characterizing the
uncertainties goes to zero, this set of quantum states converges to
the unique state consistent with the Born rule. We furthermore
establish a mathematical correlation between the sizes of the
intervals and the volume of states that are all consistent with an
imprecise measurement.
\andy{``volume of states" are not defined in the Gleason's section.}

@@

\begin{definition}[$\delta$-Determinism]\label{def:delta-deterministic} A
  QIVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}$ is
  $\delta$-deterministic if, for every event $P\in\events$, we have
  that either 
  $\bar{\mu}\left(P\right)\subseteq\left[0,\delta\right]$ or
  $\bar{\mu}\left(P\right)\subseteq\left[1-\delta,1\right]$. 
\end{definition}

\begin{table}
  \noindent \centering{}\caption{\label{tab:classical-IVPMs}Example 
    QIVPMs~$\bar{\mu}_{A}$, $\bar{\mu}_{B}$ and $\bar{\mu}_{C}$ on
    commuting events $P$.}
\begin{tabular}{cccc}
\toprule 
\addlinespace
$P$ & $\bar{\mu}_{A}(P)$ & $\bar{\mu}_{B}(P)$ & $\bar{\mu}_{C}(P)$\tabularnewline\addlinespace
\midrule
\midrule 
\addlinespace
$\mathbb{0}$ & $\imposs$ & $\imposs$ & $\imposs$\tabularnewline\addlinespace
\midrule 
\addlinespace
$\proj{0}$ & $\necess$ & $\left[\frac{2}{3},1\right]$ & $\left[\frac{2}{3},\frac{2}{3}\right]$\tabularnewline\addlinespace
\midrule 
\addlinespace
$\proj{1}$ & $\imposs$ & $\left[0,\frac{1}{3}\right]$ & $\left[\frac{1}{3},\frac{1}{3}\right]$\tabularnewline\addlinespace
\midrule 
\addlinespace
$\mathbb{1}$ & $\necess$ & $\necess$ & $\necess$\tabularnewline\addlinespace
\bottomrule
\end{tabular}
\yutsung{$\bar{\mu}_{A}$ is neither a QIVPM nor a restriction of a QIVPM
according to the current definition. Need to fix this example or fix the definition.}
\end{table}

@@

Our framework allows us to generalize the above theorem to state that
for small enough $\delta$, it is impossible to have
$\delta$-deterministic QIVPMs. We next prove a main result of this
paper which is a stronger statement of contextuality that includes the
effects of finite-precision.  Every QIVPM must map some events to
truly uncertain intervals, not just ``almost definite intervals.'' 
\andy{contextuality is never defined.}
\yutsung{Actually, the second sentence in the introduction defines
``contextuality''. The problem is that there is no direct relation between what
defined in the introduction and here.}

@@

We have thus quantified one important aspect of uncertainty in quantum
mechanics---the effect of the imprecise nature of devices---which is a novel
addition to the theory of measurement. Indeed, as Heisenberg
emphasized in his famous microscope
example~\cite{Heisenberg1983apsrev4}, the conventional theory
of measurement states that it is impossible to precisely measure any
property of a system without disturbing it somewhat. Thus, there are
fundamental limits to what one can measure and these limits have traditionally 
been attributed to complementarity. Our imprecision represents an 
\emph{additional} source of indeterminacy beyond the inherent probabilistic nature
of quantum mechanics.
\andy{Make sure Amr is OK with the ``beyond'' here. This was previously ``to''.}
% The significance of this observation is that the uncertainty in the
% intervals is, in contrast to the classical case, \emph{a fundamental
%   part of quantum theory}. There are in fact two sources of
% uncertainty: one due to the finite measurement resources and one due
% to complementarity, which requires every quantum framework to be truly
% probabilistic.

@@

Foundational concepts in quantum mechanics, such as the Kochen-Specker
and Gleason theorems, rely in subtle ways on the use of
unbounded resources. By assuming infinitely precise measurements,
these two insightful theorems form the foundations of two fundamental aspects of
quantum mechanics. On the one hand, the Kochen-Specker result reveals a
distinctive aspect of physical reality, the fact that it is
contextual, and, on the other hand, Gleason's theorem establishes a
relationship between quantum states and probabilities uniquely defined
by Born's rule.  Our goal in this paper has been to analyze the
physical consequences of a mathematical framework that allows for
finite precision measurements by introducing the concept of quantum
interval-valued probability. This framework incorporates uncertainty
in the measurement results by defining fuzzy probability measures, and
includes standard quantum measurement as the particular instance of
sharp, infinitely precise, intervals.  
\andy{``quantum measurement'' should be replaced by the }
In addition, we showed how
these two theorems emerge as limiting cases of this same framework,
thus connecting two seemingly unrelated aspects of quantum physics. We
noted that {\it arbitrary\/} finite precision measurement conditions
can nullify the main tenets of both theorems. However, by carefully
specifying experimental uncertainties, we were able to establish
rigorous bounds on the validity of these two theorems. Therefore, we
have established a context in which infinite precision quantum
mechanical theories can be reconciled with finite precision quantum
mechanical measurements, and have provided a possible resolution of
the Meyer-Mermin debate on the impact of finite precision on the
Kochen-Specker theorem~\cite{PhysRevLett.83.3751,Mermin1999}.
